---
title: "final analysis 11"
output: html_document
date: "2023-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#libraries
library(tidyverse)
library(lme4)
library(lmerTest)
library(psych)
library(MuMIn)
library(sjPlot)
library(ez)
```

The dataframe "dem" contains everything from the demographic questionnaire. "newdistance" contains the most recent SHQ data, and "merged" contains older SHQ data. "levelprops" contains the proportions of levels participants completed.
```{r}
#reading in data
newdistance <- read.csv("final.csv")
merged <- read.csv("merged_final.csv")
levelprops <- read.csv("merged_final2.csv") #data with proportions of levels played

#extracting important values and joining

newdistance <- newdistance %>% select(short_id, zscore, distance) #extracting the new distances
merged <- merged %>% select(!zscore, distance) #removing old distances

#creating the final DF
joined <- inner_join(newdistance, merged, by="short_id")


```

#Demographic Data

```{r}
gender_data_joined <- joined %>% group_by(What.gender.are.you.) %>% summarise(n = length(How.old.are.you.), mean_age = mean(How.old.are.you.), self_reported_nav = mean(How.good.are.you.at.navigating.), avg_physical_activity = mean(Considering.the.last.6.months..how.many.hours.of.physical.activity.did.you.do.per.week.on.average.), avg_alcohol_consumption = mean(How.many.units.of.alcohol.do.you.consume.per.week.on.average.), avg_caffeine = mean(How.many.cups.of.caffeinated.beverages.do.you.have.a.day.)) 

joined %>% count(How.old.are.you.)

#ethnicity information

joined %>% count(What.is.your.ethnicity.)

#years of education

educ <- joined %>% summarise(m_fe = mean(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.), sd_fe = sd(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.))

educ_gender <- joined %>% group_by(What.gender.are.you.) %>% summarise(m_fe = mean(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.), sd_fe = sd(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.)) 

#proportion of levels played

levelprops %>% count(proportionoflevelsplayed)
```

##GPS and Video Game Descriptive Statistics
Exploring patterns in GPS and Video Game questionnaire responses. 

#Video Games and Smartphone Use
```{r}
#VIDEO GAME AND SMARTPHONE QUESTIONS
#mean and sd of responses on video game and smartphone questions

videogame_smartphone_desc <- joined %>% summarise(mean_videogames = mean(How.many.hours.do.you.play.video.games.per.week.), sd_videogames = sd(How.many.hours.do.you.play.video.games.per.week.), mean_smartphone_games = mean(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.), sd_smartphone_games = sd(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.), mean_smartphone = mean(How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week.), sd_smartphone = sd(How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week.))

#Same, but grouping by Gender
videogame_smartphone_desc_gender <- joined %>% group_by(What.gender.are.you.) %>% summarise(mean_videogames = mean(How.many.hours.do.you.play.video.games.per.week.), sd_videogames = sd(How.many.hours.do.you.play.video.games.per.week.), mean_smartphone_games = mean(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.), sd_smartphone_games = sd(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.), mean_smartphone = mean(How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week.), sd_smartphone = sd(How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week.))
```

#GPS and Navigation Aids
```{r}
#mean/sd responses on every item of the GPS reliance scale

gpsreliance_desc <- joined %>% summarise(mean_gps_general = mean(How.often.would.you.use.GPS.to.navigate.in.general.), sd_gps_general = sd(How.often.would.you.use.GPS.to.navigate.in.general.), mean_gps_friends = mean(You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.), sd_gps_friends = sd(You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.), mean_gps_doctor = mean(You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.), sd_gps_doctor = sd(You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.), mean_gps_family = mean(You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.), sd_gps_family = sd(You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.), mean_newroutes = mean(You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.), sd_newroutes = sd(You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.), mean_unfamiliar_destination = mean(How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.), sd_unfamiliar_destination = sd(How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.), mean_newcity = mean(When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.), sd_newcity = sd(When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.), mean_gpsoverreliance = mean(I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary), sd_gps_overreliance = sd(I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary))

#Same, but grouping by Gender

gpsreliance_desc_gender <- dem %>% group_by(What.gender.are.you.) %>% summarise(mean_gps_general = mean(How.often.would.you.use.GPS.to.navigate.in.general.), sd_gps_general = sd(How.often.would.you.use.GPS.to.navigate.in.general.), mean_gps_friends = mean(You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.), sd_gps_friends = sd(You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.), mean_gps_doctor = mean(You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.), sd_gps_doctor = sd(You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.), mean_gps_family = mean(You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.), sd_gps_family = sd(You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.), mean_newroutes = mean(You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.), sd_newroutes = sd(You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.), mean_unfamiliar_destination = mean(How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.), sd_unfamiliar_destination = sd(How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.), mean_newcity = mean(When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.), sd_newcity = sd(When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.), mean_gpsoverreliance = mean(I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary), sd_gps_overreliance = sd(I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary))

```


#GPS Reliance Composite Measure / Means
Here we create a composite score for the GPS reliance scale by calculating participants' mean response. This also reduces the impact of influential outliers.

```{r}
#joined dataset
joined <- joined %>% mutate(GPSreliance_total = (How.often.would.you.use.GPS.to.navigate.in.general.+You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.+You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.+You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.+You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.+How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.+When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.+I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary)/7)

#same for the level proportions dataset

levelprops <- levelprops %>% mutate(GPSreliance_total = (How.often.would.you.use.GPS.to.navigate.in.general.+You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.+You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.+You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.+You.usually.travel.a.specific.route.to.go.to.your.friend.s.house..This.time..you.think.you.may.get.there.faster.by.taking.a.different.route..How.often.do.you.take.new.routes.to.travel.to.places.you.have.visited.before.+How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.+When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.+I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary)/7)

#MEAN GPS COMPOSITE SCORE (BY GENDER)

gender_GPS_total <- joined %>% group_by(What.gender.are.you.) %>% summarise(mean_gps_total = mean(GPSreliance_total), sd_gps_total = sd(GPSreliance_total))

#MEAN GPS COMPOSITE SCORE(OVERALL)

gender_GPS_overall <- joined %>% summarise(mean_gps_total = mean(GPSreliance_total), sd_gps_total = sd(GPSreliance_total))
```

##Navigation Distance Statistics
Looking at patterns in navigation distances. In this dataset, navigation distances were taken from levels 1 and 11

#Navigation Distance by Gender / Overall
```{r}
gender_navdistance <- joined %>% group_by(What.gender.are.you.) %>% summarise(mean_distance = mean(distance.x), sd_distance = sd(distance.x))

#generally

joined %>% summarise(md = mean(distance.x), sdd = sd(distance.x))
```

#Navigation Distance by Video Game Genre
```{r}
gamegenre_navdistance <- joined %>% group_by(What.genre.of.video.game.do.you.most.commonly.play.) %>% summarise(m_navdistance = mean(distance.x), sd_navdistance=sd(distance.x)) %>% mutate(What.genre.of.video.game.do.you.most.commonly.play. = recode(What.genre.of.video.game.do.you.most.commonly.play., "0"='action',"1"='adventure',"2"='FPS',"3"='Roleplaying',"4"='Simulation',"5"='Strategy', "6"='Sport',"7"="Other","8"='Does not play video games'))

joined %>% count(What.genre.of.video.game.do.you.most.commonly.play.) 
```

##FIGURES
Any and all figures used in the report and analysis. Remember, higher navigation distance = inferior performance.

#Video Game / Smartphone Use Figures

```{r}
#Relationship Between Video Game Experience (general) and Smartphone Use

ggplot(dem, aes(x = How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week., y = How.many.hours.do.you.play.video.games.per.week.))+
  geom_point(shape = "triangle")+
  geom_smooth(method="lm", se=F, color="darkgrey")+
  theme_classic()+
  labs(x = "Hours of smartphone/tablet use per week", y = "Video Game hours (general) per week")

```


```{r}
#Relationship Between Video Game Experience (general) and Video Game Experience (smartphone)

ggplot(joined, aes(x = How.many.hours.do.you.play.video.games.per.week., y = How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.))+
  geom_point(shape = "triangle")+
  geom_smooth(method="lm", se=F, color="darkgrey")+
  theme_classic()+
  labs(x = "Video Game Hours (general) Per Week", y = "Video Game Hours (smartphone) per week")
```


```{r}
#Relationship Between Video Game Experience (General) and Navigation Distance

ggplot(joined, aes(x = distance.x, y = How.many.hours.do.you.play.video.games.per.week.))+
  geom_point(shape="triangle")+
  geom_smooth(method="lm", se=F, color="darkgrey")+
  theme_classic()+
  labs(x = "Video Game Hours Per Week", y = "Navigation Distance")
```


```{r}
#Relationship Between Smartphone Use and Navigation Distance

ggplot(joined, aes(x = How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week., y = distance.x))+
  geom_point(shape="triangle")+
  geom_smooth(method="lm", se=F, color="darkgrey")+
  theme_classic()+
  labs(x = "Smartphone Use (hours)", y = "Navigation Distance")
```

#GPS Figures

```{r}
#Relationship Between GPS Reliance total and Navigation Distance

ggplot(joined, aes(x = GPSreliance_total, y = distance.x))+
  geom_point(shape="triangle")+
  geom_smooth(method="lm", se=F, color="darkgrey")+
  theme_classic()+
  labs(x = "GPS Reliance", y = "Navigation Distance")
```


```{r}
#Relationship Between GPS Reliance Questionnaire Items (Individually) and Navigation Distance


#first creating a subset with the 7 GPS reliance items and changing to long format
GPS_reliance_long <- joined %>% select(contains("GPS"), distance.x) %>% rename(q1=1,q2=2,q3=3,q4=4,q5=5,q6=6,q7=7) %>% select(!8) %>% gather(Question, Response, q1:q7)

#plotting

ggplot(GPS_reliance_long, aes(x = Response, y = distance.x, color = Question))+
  geom_point(shape="triangle")+
  geom_smooth(method="lm", se=F)+
  theme_minimal()+
  labs(y = "Navigation Distance", x = "Response", color = "GPS Reliance Scale Item")
```
q1 = how often would you use GPS to navigate in general?
q2 = you are meeting friends at a new restaurant, and you are travelling there for the first time - how often would you use GPS?
q3 = (doctor scenario) you have to travel a new route to reach a destination you have visited before, how often would you use a GPS?
q4 = You are leaving home to visit family and are taking the same route as always. How often would you use GPS?
q5 = How often do you use GPS to travel outside of the area where you live?
q6 = When finding your way around a city outside of your hometown, how often would you use a GPS?
q7 = I overrely on GPS in situations where it is not really necessary.

#INFERENTIAL STATISTICS - Video Games & Smartphone
Here we are conducting statistical tests assessing various relationships in the data.

```{r}
#Pearson Correlation Between Smartphone Use and Navigation Distance

cor.test(joined %>% pull(How.many.hours.per.week.do.you.use.a.smartphone.or.a.tablet.per.week.),
         joined %>% pull(distance.x), method="pearson")

#R=0.31, P=0.03*
```

```{r}
#Pearson Correlation Between Video Game Experience and Video Game Experience on smartphones

cor.test(joined %>% pull(How.many.hours.do.you.play.video.games.per.week.),
         joined %>% pull(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.), method="pearson")

#R = 0.704, p<0.001***
```


```{r}
#Pearson Correlation Between Video Game Experience and Navigation Distance

cor.test(joined %>% pull(How.many.hours.do.you.play.video.games.per.week.),
         joined %>% pull(distance.x), method="pearson")



#R=-0.03, P=0.8
```

```{r}
#Pearson Correlation Between Video Game Experience (on smartphones) and Navigation Distance

cor.test(joined %>% pull(How.many.hours.do.you.play.video.games.on.a.smartphone.or.a.tablet.per.week.),
         joined %>% pull(distance.x), method="pearson")

#R-0.08, P=0.61
```

```{r}
ezANOVA(joined, distance.x, wid=short_id, between=What.genre.of.video.game.do.you.most.commonly.play.)
```


#INFERENTIAL STATISTICS - DISTANCE MEASURES + DEMOGRAPHICS

```{r}
#Gender Difference in Navigation Distance

t.test(joined %>% filter(What.gender.are.you.==1) %>% pull(distance.x),
       joined %>% filter(What.gender.are.you.==2) %>% pull(distance.x))

#t(29) = 1.06, p=0.29, not significant

#Gender Difference in Years of Education

t.test(joined %>% filter(What.gender.are.you.==1) %>% pull(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.),
       joined %>% filter(What.gender.are.you.==2) %>% pull(How.many.years.of.formal.education.do.you.have..starting.from.kindergarten..K5..or.age.5.))

#t(10)=1.15, p=0.27
```


#INFERENTIAL STATISTICS - GPS

```{r}
#Pearson Correlation Between GPS Reliance Total and Navigation Distance

cor.test(joined %>% pull(GPSreliance_total),
         joined %>% pull(distance.x), method="pearson")

#R = -0.03, P=0.84
```


```{r}
#Multiple Regression - Distance ~ GPS Reliance Items

mod1 <- lm(distance.x ~ How.often.would.you.use.GPS.to.navigate.in.general.+You.are.meeting.nfriends.at.a.new.restaurant..and.you.are.traveling.there.nfor.the.first.time..How.often.do.you.use.a.GPS.in.such.a.nsituation.+You.usually.leave.from.home.to.go.to.a.doctor.s.appointment..This.time..however..your.appointment.is.scheduled.right.after.work..Therefore..you.have.to.travel.a.new.route.to.get.to.a.destination.you.have.visited.before..How.often.do.you.use.a.GPS.in.such.a.situation.+You.usually.leave.from.home.to.visit.your.family..You.are.taking.the.same.route.as.you.always.do..How.often.do.you.use.a.GPS.in.such.a.situation.+How.often.do.you.use.a.GPS.to.travel.to.a.destination.outside.of.the.area.where.you.live.+When.finding.your.way.around.a.city.outside.of.your.hometown..how.often.do.you.use.a.GPS.+I.over.rely.on.using.GPS.in.situations.where.it.is.not.really.necessary, data=joined)

summary(mod1) #no significant effects
tab_model(mod1, file = "DistanceGPSregression.doc")


```

#INFERENTIAL STATISTICS - INTERACTION + GENDER RANDOM EFFECT
Linear Mixed Effects Model to assess the relationship between video game experience, GPS reliance, and gender.

```{r}
joined <- joined %>% mutate(What.gender.are.you. = as.factor(What.gender.are.you.))

#fitting model
mod_full <- lmer(distance.x ~ How.many.hours.do.you.play.video.games.per.week.*GPSreliance_total + (1 | What.gender.are.you.), data=joined)
summary(mod_full)
tab_model(mod_full, file = "LMERinteraction.doc")
r.squaredGLMM(mod_full)
```